---
title: 'Reproducible Research: Part 1'
author: "Ribery"
date: "9 октомври 2015 г."
output: html_document
---

Part 1



1. Load the data (i.e. read.csv())

```{r, echo=FALSE}
# clear memory
rm(list=ls())

# Load the raw activity data
activity <- read.csv("C:/RBook/Reproducible Research/1/activity.csv", stringsAsFactors=FALSE)

head(activity)

library(lattice)
activity$date <- as.Date(activity$date, "%Y-%m-%d")

```



What is the mean total number of steps taken per day?
For this part of the assignment, you can ignore the missing values in the dataset.

Make a histogram of the total number of steps taken each day


```{r, echo=FALSE}

StepsTotal <- aggregate(steps ~ date, data = activity, sum, na.rm = TRUE)

hist(StepsTotal$steps, main = "Total steps by day", xlab = "day", col = "red")

```


And the mean and median and mean is
```{r, echo=FALSE}
median(StepsTotal$steps)
mean(StepsTotal$steps)

```


The second approach is to make a data frame first with the values that I need I create a data frame with the days and the total of steps by day.

```{r, echo=FALSE}
steps <- rep(NA, 61)
day <- rep("NA", 61)
stepsday <- tapply(activity$steps, activity$date, sum, na.rm = T)
length(stepsday)

```



```{r, echo=FALSE}
for (i in 1:61) {
    steps[i] <- stepsday[[i]]
    day[i] <- names(stepsday)[i]
}


```


Then the data frame is

```{r, echo=FALSE}
df <- data.frame(day, steps)
head(df)

```



Now i make the histogram

```{r, echo=FALSE}
hist(df$steps, main = "Total steps by day", xlab = "day", col = "yellow")


```


What is the average daily activity pattern?

Make a time series plot (i.e. type = УlФ) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

Now I get mean of steps and time series plot

```{r, echo=FALSE}
time_series <- tapply(activity$steps, activity$interval, mean, na.rm = TRUE)
plot(row.names(time_series), time_series, type = "l", xlab = "5-min interval", 
    ylab = "Average across all Days", main = "Average number of steps taken", 
    col = "blue")
```


Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r, echo=FALSE}
max_interval <- which.max(time_series)
names(max_interval)

```


Imputing missing values

Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r, echo=FALSE}
activity_NA <- sum(is.na(activity))
activity_NA

```




Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval,etc

Fist Na replaced by mean in 5 min interval

```{r, echo=FALSE}
StepsAverage <- aggregate(steps ~ interval, data = activity, FUN = mean)
fillNA <- numeric()
for (i in 1:nrow(activity)) {
    obs <- activity[i, ]
    if (is.na(obs$steps)) {
        steps <- subset(StepsAverage, interval == obs$interval)$steps
    } else {
        steps <- obs$steps
    }
    fillNA <- c(fillNA, steps)
}
```


Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r, echo=FALSE}
new_activity <- activity
new_activity$steps <- fillNA


```


Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

```{r, echo=FALSE}
StepsTotal2 <- aggregate(steps ~ date, data = new_activity, sum, na.rm = TRUE)
hist(StepsTotal2$steps, main = "Total steps by day", xlab = "day", col = "blue")
```


And the mean and median is

```{r, echo=FALSE}
mean(StepsTotal2$steps)
median(StepsTotal2$steps)

```



Are there differences in activity patterns between weekdays and weekends?

For this part the weekdays() function may be of some help here. Use the dataset with the filled-in missing values for this part.

Create a new factor variable in the dataset with two levels weekday and weekend indicating whether a given date is a weekday or weekend day.

```{r, echo=FALSE}

day <- weekdays(activity$date)
daylevel <- vector()
for (i in 1:nrow(activity)) {
    if (day[i] == "Saturday") {
        daylevel[i] <- "Weekend"
    } else if (day[i] == "Sunday") {
        daylevel[i] <- "Weekend"
    } else {
        daylevel[i] <- "Weekday"
    }
}
activity$daylevel <- daylevel
activity$daylevel <- factor(activity$daylevel)

stepsByDay <- aggregate(steps ~ interval + daylevel, data = activity, mean)
names(stepsByDay) <- c("interval", "daylevel", "steps")

```



Make a panel plot containing a time series plot (i.e. type = УlФ) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). The plot should look something like the following, which was creating using simulated data.

```{r, echo=FALSE}
xyplot(steps ~ interval | daylevel, stepsByDay, type = "l", layout = c(1, 2), 
    xlab = "Interval", ylab = "Number of steps")

```





